---
title: "P8105_HW1_lz2950"
author: "Lehan Zou"
date: "2023-09-21"
output: github_document
---

### load necessary packages
```{r}
library("moderndive")
library("tidyverse")
library("ggplot2")
library("tidyverse")
```
## Problem 1
```{r}
#load dataset "early_january_weather"
data("early_january_weather")
summary(early_january_weather)

```

This data set contains 15 variables: `r variable.names(early_january_weather)`. Since humidity and temperature are going to be specified below, I consider those two to be important variables. For temperature, the `min`, `max`, `median` values are `r min(early_january_weather$temp)`, `r max(early_january_weather$temp)`, `r median(early_january_weather$temp)`; For humidity,  the `min`, `max`, `median` values are `r min(early_january_weather$humid)`, `r max(early_january_weather$humid)`, `r median(early_january_weather$humid)`. 
 

* size of the data set
The data set has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.

* mean temperature
The mean temperature of this data set is `mean(early_january_weather$temp)`.

* plot
```{r}
#using ggplot to make scatter plot, define x,y variables and color
ggplot(early_january_weather, aes(time_hour, temp, color = humid)) + 
  geom_point()+
  labs(title = "Scatterplot of Temperature ~ Time")
#save image
ggsave("hw1_scatterplot.png")
```
From the scatter plot, the temperature fluctuates within days but the overall trend is increasing.The humidity is relatively steady but increases between Jan 12 to Jan 14. And it drop back on Jan 14. Also matches the highest temperature over this time period. 

### Problem 2
Create a data frame comprised of:  

 * a random sample of size 10 from a standard Normal distribution  
 * a logical vector indicating whether elements of the sample are greater than 0  
 * a character vector consisting of the sample converted to character  
 * a factor vector measuring the value of the sample

```{r}
#define vectors
random_sample <- rnorm(10)
logical_vec <- random_sample > 0
character_vec <- as.character(random_sample)
factor_vec <- factor(rep(c("Level 1", "Level 2", "Level 3"), length.out = 10)) 
```

```{r}
#Combine those vectors to a data frame
data_frame <- data.frame(RandomSample = random_sample, LogicalVector = logical_vec, CharacterVector = character_vec, FactorVector = factor_vec)
#show the data frame
data_frame
```
* calculate mean


```{r}
for(variable in variable.names(data_frame))
  print(paste("The mean for variable '", variable, "' is ", data_frame %>% pull(variable) %>% mean, ".", sep = ""))
```
This shows it works for the numeric and logical vectors while character vector and factor vector are NA values.

```{r}
# Convert logical, character, and factor variables to numeric
numeric_logical <- as.numeric(data_frame$LogicalVector)
numeric_logical

numeric_character <- as.numeric(data_frame$CharacterVector)
numeric_character

numeric_factor <- as.numeric(data_frame$FactorVector)
numeric_factor
```
All of the vectors got the mean result after convert to numerical variables. This indicate for mean() function, logical vector can be converted by default while character vector and factor vector are not.  